---
title: "Eden"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown


For your RScript:
```{r}
#setwd("C:/Users/User/Desktop/Titanic")
```


For rmarkdown:
```{r setup}
#knitr::opts_knit$set(root.dir = ("C:\Users\User\Desktop\Titanic")
getwd()

```



## Load the'test' and 'train' sets - treat empty values as Na's 

```{r}
setwd("C:/Users/User/Desktop/Titanic")
train <- read.csv("train.csv",na.strings = "")
test<-read.csv("test.csv",na.strings = "")
```

##  Data Clening

1. 1. Turning the survived and Pclass values into factors features - train setst set

```{r}
   train$Survived <- as.factor(train$Survived)
   train$Pclass <- as.factor(train$Pclass)
```

2. Adding 'Survived'column to Test Set and turning the survived and Pclass values into factors features - test set
```{r}
  test$survived <- NA 
  test$survived <- as.factor(test$survived)
  test$Pclass <- as.factor(test$Pclass)

```

3. PassengerId won't help us much predicting the paseenger's destiny so we can take this property off and  we can see that the Name column is currntly not being helpful, but we can at least extarct 
the title from it. 
There are quite a few titles but I would like to reduce them all to Mrs, Miss Mr and Master.To do so I will need to use a fuction which srerches for substrings. Grab title from passenger Name

```{r}

  train<-subset(train,select = -c(1))
  train$Title <- gsub('(.*, )|(\\..*)', '', train$Name)
  test$Title <- gsub('(.*, )|(\\..*)', '', test$Name)
  table(train$Sex, train$Title)
  table(test$Sex,test$Title)
  otherTitle <-c ('Capt','Col','Don','Dr','Jonkheer','Lady','Major','Rev','Sir','the Countess','Dona','Dr')
  train$Title[train$Title == 'Mlle'] <- 'Miss'
  test$Title[test$Title == 'Mlle'] <- 'Miss'
  train$Title[train$Title == 'Ms'] <- 'Miss'
  test$Title[test$Title == 'Ms'] <- 'Miss'
  train$Title[train$Title == 'Mme'] <- 'Mrs'
  test$Title[test$Title == 'Mme'] <- 'Mrs'
  # apply 'other' to other Title 
  train$Title[train$Title %in% otherTitle] <- "other"
  test$Title[test$Title %in% otherTitle] <- "other"
  # Check the title again
  table(train$Sex, train$Title)
  table(test$Sex, test$Title)
 

```
4. Now, we can check the family size, We will put it a new column 'FamilySize'
We will calculate the size that way > size = SibSp+Parch +1 (the passenger)s

```{r}
train$FamilySize <- train$SibSp + train$Parch + 1
test$FamilySize <- test$SibSp + test$Parch + 1

# let's discretize family size
train$DisSize[train$FamilySize == 1] <- 'alone'
train$DisSize[train$FamilySize < 4 & train$FamilySize >1] <- 'small'
train$DisSize[train$FamilySize > 3 ] <- 'big'

test$DisSize[test$FamilySize==1] <- 'alone'
test$DisSize[test$FamilySize < 4 & test$FamilySize >1] <- 'small'
test$DisSize[test$FamilySize > 3 ] <- 'big'

# let's take a look at the Fare column and divide it into 4 groups 
train$FClass[train$Fare <= 10] <- 'Low'
train$FClass[train$Fare <= 40 & train$Fare > 10] <- 'Noraml'
train$FClass[train$Fare <= 100 & train$Fare > 40] <- 'Med'
train$FClass[train$Fare > 100] <- 'High'
test$FClass[test$Fare <= 10] <- 'Low'
test$FClass[test$Fare <= 40 & test$Fare > 10] <- 'Noraml'
test$FClass[test$Fare <= 100 & test$Fare > 40] <- 'Med'
test$FClass[test$Fare > 100] <- 'High'

```

5. We are going to handle the 'Age'column 

```{r}
library(rpart)
train<-subset(train,select = -c(3,8,10))
test<-subset(test,select = -c(3,8,10))

train$sibSpDis <- sapply(train$SibSp, FUN=function(x) {if(x >=3) x ='flag' else x=x})
test$sibSpDis <- sapply(test$SibSp, FUN=function(x) {if(x >=3) x ='flag' else x=x})

train$sibSpDis<- as.factor(train$sibSpDis)
test$sibSpDis<- as.factor(test$sibSpDis)


AgeReplace <- rpart(Age~ Pclass + Sex + FamilySize + Title + Parch + Embarked + DisSize + FClass + sibSpDis, data=train[!is.na(train$Age),],method="anova")
AgePredict <- predict(AgeReplace, train[is.na(train$Age),])
train$Age[is.na(train$Age)] <- AgePredict

AgeReplace <- rpart(Age ~ Pclass + Sex + FamilySize + Title + Parch + Embarked + DisSize + FClass + sibSpDis, data=test[!is.na(test$Age),],method="anova")
AgePredict <- predict(AgeReplace, test[is.na(test$Age),])
test$Age[is.na(test$Age)] <- AgePredict

```


##Random Forest 


```{r}
#install.packages("party")
#install.packages("caret")
library(party)
library(caret)
library(randomForest)


set.seed(123)

train$Age[is.na(train$Age)] <- -1
train$Fare[is.na(train$Fare)] <- median(train$Fare, na.rm=TRUE)
train$Embarked[is.na(train$Embarked)] = "S"

test$Age[is.na(test$Age)] <- -1
test$Fare[is.na(test$Fare)] <- median(test$Fare, na.rm=TRUE)
test$Embarked[is.na(test$Embarked)] = "S"

fit <- randomForest(as.factor(Survived) ~ Pclass + Sex + FamilySize + Title + Parch + Embarked + DisSize + FClass + sibSpDis,
                      data=train, 
                      importance=TRUE, 
                      ntree=2000)

varImpPlot(fit)
Prediction <- predict(fit, test)
submit <- data.frame(PassengerId = test$PassengerId, Survived = Prediction)
write.csv(submit, file = "C:/Users/User/Desktop/Titanic/firstforest.csv", row.names = FALSE)
```

## RPart

```{r}
library(rpart)
tree <- rpart(Survived ~ Pclass + Sex + FamilySize + Title + Parch + Embarked + DisSize + FClass + sibSpDis,
              data=train,
              method="class")

Prediction <- predict(tree, test, type = "class")
submit <- data.frame(PassengerId = test$PassengerId, Survived = Prediction)
 write.csv(submit, file = "C:/Users/User/Desktop/Titanic/firsTree", row.names = FALSE)
 
```

C5.0

```{r}
indices = sample(1:nrow(train),nrow(train)*0.75)

#install.packages(C50)
library(C50)
set.seed(123)
ToTrain<- train[indices,]
ToTest<- train[-indices,]
C50 <-C5.0(Survived  ~ Pclass + Sex + FamilySize + Title + Parch + Embarked + DisSize + FClass + sibSpDis , data=train)
Prediction <- predict(C50,test)
submit <- data.frame(PassengerId = test$PassengerId, Survived = Prediction)
write.csv(toSubmit, file = "C:/Users/User/Desktop/Titanic/firstC5.0", row.names = FALSE)


```



